---
title: "Chapter 5 Problem Set Solutions"
output:
  html_document:
    df_print: paged
---

```{r packages}
library(readxl)
setwd("C:/Users/timma/Documents/GitHub/FIN6572Fall2020/Problem Sets/Ch5")
```

### 5.1

a. $\Pr(X<325) = \Pr\left(Z<\dfrac{325-330}{4}\right)$
```{r 5.1a}
pnorm(325,330,4)
```
b. $\Pr\left(\bar{X}<325| N=6\right)=\Pr\left(Z<\dfrac{325-330}{4/\sqrt{6}}\right)$
```{r 5.1b}
pnorm(325,330,4/sqrt(6))
```
c. $\Pr\left(\bar{X}<325| N=6\right)=\Pr\left(Z<\dfrac{325-330}{4/\sqrt{12}}\right)$
```{r 5.1c}
pnorm(325,330,4/sqrt(12))
```


### 5.2

$\Pr\left(\bar{X}>200\right)=\Pr\left(Z>\dfrac{200-180}{28/\sqrt{16}}\right)$
```{r 5.2}
1-pnorm(200, 180, 28/sqrt(16))
```


### 5.3

a. $\Pr\left(X>35\right)=\Pr\left(Z>\dfrac{35-33.8}{3.5}\right)$
```{r 5.3a}
1-pnorm(35,33.8,3.5,1)
```

b. $\Pr\left(\bar{X}>35\right)=\Pr\left(Z>\dfrac{35-33.8}{3.5\sqrt(4)}\right)$
```{r 5.3b}
1-pnorm(35,33.8,3.5/sqrt(4))
```

c. $\left(\Pr\left(X>35\right)\right)^4$
```{r 5.3c}
(1-pnorm(35,33.8,3.5))^4
```


### 5.4

a. $\mathbb{E}\left[\hat{P}\right] = 1/6 \approx 0.17 = p$
$se\left(\hat{p}\right) = \sqrt{\dfrac{0.17(1-0.17)}{200}}$
```{r 5.4a}
se = sqrt(((1/6)*(5/6))/200)
se
```

$\Pr\left(\bar{P}>20\right) = \Pr\left(Z>\dfrac{0.20-0.17}{\sqrt{\frac{0.17(1-0.17)}{200}}}\right)$

Note: This is slightly off from the connect solution because why round when you don't have to.
```{r 5.4b}
1-pnorm(0.2,1/6,se)
```


### 5.5

```{r 5.5data}
myData = read_xlsx("Ch05_Q17_Data_File.xlsx")
```

A 90% confidence interval corresponds with an upper tail of 5% (0.95 in qt function gets this 95th percentile).
```{r 5.5a}
Error = qt(0.95,5) * sd(myData$Customers)/sqrt(6)
Error
```

Construct 90% confidence interval from mean and margin of error.
```{r 5.5b}
Lower = mean(myData$Customers) - Error
Upper = mean(myData$Customers) + Error
c(Lower,Upper)
```

Alternatively, one can run a one-sample t-test.
```{r 5.5alt}
t.test(myData$Customers, conf.level=0.90)
```


### 5.6

```{r 5.6data}
myData = read_xlsx("Ch05_Q25_Data_File.xlsx")
```


```{r 5.6}
t.test(myData$Debt, conf.level=0.90)
t.test(myData$Debt, conf.level=0.95)
sd(myData$Debt)
```

### 5.7

For 95% confidence interval, upper tail is 97.5th quantile, which is then scaled by standard deviation of the estimate.
```{r 5.7a}
moe = qt(0.975,1025) * sqrt((0.47*(1-0.47))/1026)
moe
```

```{r 5.7b}
lb = 0.47 - moe
ub = 0.47 + moe
c(lb,ub)
```


### 5.9

Calculate t-statistic:
```{r 5.9b}
t = (2.8-3)*sqrt(45)
t
```

p-value:
```{r 5.9c}
pt(t,44)
```


### 5.10

```{r 5.10data}
myData = read_xlsx("Ch05_Q47__Data_File.xlsx")
```

Calculate t-statstic:
```{r 5.10c}
mu = mean(myData$Debt)
s = sd(myData$Debt)
t = (900-mu)/(s/sqrt(length(myData$Debt)))
t
```



